### 2023-03-13 paper/paper.tex
Add draft paper

% Template for ICIP-2022 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{Learned image compression}
%
% Single address.
% ---------------
\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
\address{Author Affiliation(s)}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
The abstract should appear at the top of the left-hand column of text, about
0.5 inch (12 mm) below the title area and no more than 3.125 inches (80 mm) in
length.  Leave a 0.5 inch (12 mm) space between the end of the abstract and the
beginning of the main text.  The abstract should contain about 100 to 150
words, and should be identical to the abstract text submitted electronically
along with the paper cover sheet.  All manuscripts must be in English, printed
in black ink.
\end{abstract}
%
\begin{keywords}
One, two, three, four, five
\end{keywords}
%
\section{Introduction} % Juha
\label{sec:intro}
With the rise of cloud services and mobile devices, the amount of stored digital media has grown rapidly. 
Large part of this data is stored images which take considerable amount of space.
Traditionally these images have been compressed or encoded to formats like jpeg or png which compress the images and usually lose some details (lossy compression).
With the rise of neural networks and machine learning there has been many attempts to achieve higher compression ratios and lose less details.
Many of these methods have achieved higher compression ratio with better perceived image quality and less visible artifacts but have been prohibitively expensive in terms of computational cost.

This work tries to replicate paper from Dailan et al\cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
The model is based on multiple convolutional neural networks which encode the image into latent space.
This latent representation is then passed into autoencoder which compresses it and is sent to the receiver. 
The autoencoder takes as parameters input from a model they call Space-Channel ConTeXt (SCCTX).
This model optimizes the arithmetic encoders performance with input parameters from ... (some technical stuff about scctx) TEEMU?

The output of the arithmetic decoder is passed into otherwise identical neural network as the first one in the pipeline, but in reversed order of layers with transposed convolutional layers instead of convolutional layers.
The output of this network is now our final image output.

There are still some limitations to this approach. 
One of which is that the network input image size is limited.
This means that the network cannot take as input arbitrarily sized images but of course these images can be padded or split or cropped and inputted to the network in multiple pieces.
Second limitation of this network is the amount of time it takes to encode an image and decode an image. 
Even when it is considerably faster than other learned image compression methods, in our testing it still very slow and instead of highlighted 100 microseconds for a image which size was not specified, when we tried to encode and compress an 1024x2048 sized image it took approximately 4 seconds for encoding and 5 seconds for decoding.
Most likely our implementation has many problems which would cause this value to be far away from the highlighted one, but we are about 4 magnitudes away from their highlighted value which at least to us raises some questions.

\section{Experiments}
\label{sec:experiments}
\subsection{Training} % Juha
\label{sec:training}

Training of the model was done with 6000 largest images from ImageNet \cite{imagenet} which were randomly cropped into size of 16x16 pixels.
These images were then run through the forward method of the network.
The forward method of the network does not fully compress the images but it transforms them into the latent spaces and the loss function takes into account the likelihood of good compression ratio as well as mse loss.

The model was trained with multiple different configurations. We tried $\lambda \in \{0.0008, 0,001\}$ where latent space channels $M$ and $N$ were $M=N=192$ and $\lambda \in \{0.0004, 0.0008\}$ where latent space channels $M$ and $N$ were $M=192, N=128$.
All of these models were trained for approximately 300 epochs but they could be still trained further.
Unfortunately due to lack of computational capabilities we could not train more networks and even these networks most likely could have been trained for longer but with epochs taking from 180 seconds (for smaller latent space) to 300 seconds (larger latent space), we could not train them further due to time constraints.

\vfill\pagebreak
\vfill\pagebreak

\section{Formatting your paper}
\label{sec:format}

All printed material, including text, illustrations, and charts, must be kept
within a print area of 7 inches (178 mm) wide by 9 inches (229 mm) high. Do
not write or print anything outside the print area. The top margin must be 1
inch (25 mm), except for the title page, and the left margin must be 0.75 inch
(19 mm).  All {\it text} must be in a two-column format. Columns are to be 3.39
inches (86 mm) wide, with a 0.24 inch (6 mm) space between them. Text must be
fully justified.

\section{PAGE TITLE SECTION}
\label{sec:pagestyle}

The paper title (on the first page) should begin 1.38 inches (35 mm) from the
top edge of the page, centered, completely capitalized, and in Times 14-point,
boldface type.  The authors' name(s) and affiliation(s) appear below the title
in capital and lower case letters.  Papers with multiple authors and
affiliations may require two or more lines for this information. Please note
that papers should not be submitted blind; include the authors' names on the
PDF.

\section{TYPE-STYLE AND FONTS}
\label{sec:typestyle}

To achieve the best rendering both in printed proceedings and electronic proceedings, we
strongly encourage you to use Times-Roman font.  In addition, this will give
the proceedings a more uniform look.  Use a font that is no smaller than nine
point type throughout the paper, including figure captions.

In nine point type font, capital letters are 2 mm high.  {\bf If you use the
smallest point size, there should be no more than 3.2 lines/cm (8 lines/inch)
vertically.}  This is a minimum spacing; 2.75 lines/cm (7 lines/inch) will make
the paper much more readable.  Larger type sizes require correspondingly larger
vertical spacing.  Please do not double-space your paper.  TrueType or
Postscript Type 1 fonts are preferred.

The first paragraph in each section should not be indented, but all the
following paragraphs within the section should be indented as these paragraphs
demonstrate.

\section{MAJOR HEADINGS}
\label{sec:majhead}

Major headings, for example, "1. Introduction", should appear in all capital
letters, bold face if possible, centered in the column, with one blank line
before, and one blank line after. Use a period (".") after the heading number,
not a colon.

\subsection{Subheadings}
\label{ssec:subhead}

Subheadings should appear in lower case (initial word capitalized) in
boldface.  They should start at the left margin on a separate line.

\subsubsection{Sub-subheadings}
\label{sssec:subsubhead}

Sub-subheadings, as in this paragraph, are discouraged. However, if you
must use them, they should appear in lower case (initial word
capitalized) and start at the left margin on a separate line, with paragraph
text beginning on the following line.  They should be in italics.

\section{PRINTING YOUR PAPER}
\label{sec:print}

Print your properly formatted text on high-quality, 8.5 x 11-inch white printer
paper. A4 paper is also acceptable, but please leave the extra 0.5 inch (12 mm)
empty at the BOTTOM of the page and follow the top and left margins as
specified.  If the last page of your paper is only partially filled, arrange
the columns so that they are evenly balanced if possible, rather than having
one long column.

In LaTeX, to start a new column (but not a new page) and help balance the
last-page column lengths, you can use the command ``$\backslash$pagebreak'' as
demonstrated on this page (see the LaTeX source below).

\section{PAGE NUMBERING}
\label{sec:page}

Please do {\bf not} paginate your paper.  Page numbers, session numbers, and
conference identification will be inserted when the paper is included in the
proceedings.

\section{ILLUSTRATIONS, GRAPHS, AND PHOTOGRAPHS}
\label{sec:illust}

Illustrations must appear within the designated margins.  They may span the two
columns.  If possible, position illustrations at the top of columns, rather
than in the middle or at the bottom.  Caption and number every illustration.
All halftone illustrations must be clear black and white prints.  Colors may be
used, but they should be selected so as to be readable when printed on a
black-only printer.

Since there are many ways, often incompatible, of including images (e.g., with
experimental results) in a LaTeX document, below is an example of how to do
this \cite{Lamp86}.

\section{FOOTNOTES}
\label{sec:foot}

Use footnotes sparingly (or not at all!) and place them at the bottom of the
column on the page on which they are referenced. Use Times 9-point type,
single-spaced. To help your readers, avoid using footnotes altogether and
include necessary peripheral observations in the text (within parentheses, if
you prefer, as in this sentence).

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
\begin{figure}[htb]

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{image1}}
%  \vspace{2.0cm}
  \centerline{(a) Result 1}\medskip
\end{minipage}
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{image3}}
%  \vspace{1.5cm}
  \centerline{(b) Results 3}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{image4}}
%  \vspace{1.5cm}
  \centerline{(c) Result 4}\medskip
\end{minipage}
%
\caption{Example of placing a figure with experimental results.}
\label{fig:res}
%
\end{figure}

% To start a new column (but not a new page) and help balance the last-page
% column length use \vfill\pagebreak.
% -------------------------------------------------------------------------
%\vfill
%\pagebreak

\section{COPYRIGHT FORMS}
\label{sec:copyright}

You must submit your fully completed, signed IEEE electronic copyright release
form when you submit your paper. We {\bf must} have this form before your paper
can be published in the proceedings.

\section{RELATION TO PRIOR WORK}
\label{sec:prior}

The text of the paper should contain discussions on how the paper's
contributions are related to prior work in the field. It is important
to put new work in  context, to give credit to foundational work, and
to provide details associated with the previous work that have appeared
in the literature. This discussion may be a separate, numbered section
or it may appear elsewhere in the body of the manuscript, but it must
be present.

You should differentiate what is new and how your work expands on
or takes a different path from the prior studies. An example might
read something to the effect: "The work presented here has focused
on the formulation of the ABC algorithm, which takes advantage of
non-uniform time-frequency domain analysis of data. The work by
Smith and Cohen \cite{Lamp86} considers only fixed time-domain analysis and
the work by Jones et al \cite{C2} takes a different approach based on
fixed frequency partitioning. While the present study is related
to recent approaches in time-frequency analysis [3-5], it capitalizes
on a new feature space, which was not considered in these earlier
studies."

\vfill\pagebreak

\section{REFERENCES}
\label{sec:refs}

List and number all bibliographical references at the end of the
paper. The references can be numbered in alphabetic order or in
order of appearance in the document. When referring to them in
the text, type the corresponding reference number in square
brackets as shown at the end of this sentence \cite{C2}. An
additional final page (the fifth page, in most cases) is
allowed, but must contain only references to the prior
literature.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{strings,refs}

\end{document}

### 2023-03-13 eval_mode.py
sss

@@ -155,21 +155,20 @@ def inference(model, test_dataloader, criterion, device, loss, bpp_loss, mse_los
             
             # iterate over the batch and plot the images
             img = transforms.ToPILImage()(x.squeeze().cpu())
-            rec_net = transforms.ToPILImage()(out_net['x_hat'].squeeze().cpu())
+            #rec_net = transforms.ToPILImage()(out_net['x_hat'].squeeze().cpu())
             whole = transforms.ToPILImage()(out_dec['x_hat'].squeeze().cpu())
             
-            imgs = [img, rec_net, whole]
-            titles = ['Original', 'Forward pass', 'Compression + Decompression']
+            imgs = [img, whole]#[img, rec_net, whole]
+            titles = ['Original','Compression + Decompression']#['Original', 'Forward pass', 'Compression + Decompression']
             # save with suplot iteratively
             _, axs = plt.subplots(1, len(imgs), figsize=(10, 5))
-            plt.suptitle(f'ELIC Result')
+            plt.suptitle(f'avg. bits per pixel: {bpp:.4f}')
             for ix,_ in enumerate(imgs):
                 axs[ix].imshow(imgs[ix])
                 axs[ix].set_title(titles[ix])
                 axs[ix].axis('off')
             plt.tight_layout()
             plt.savefig(f'./{output_dir}/{offset+i}.png')
-                        
 
             out_criterion = criterion(out_net, x)
             aux_loss.update(model.aux_loss())
@@ -420,7 +419,7 @@ def main(argv):
     test_dataset = ImageFolder(args.dataset, split="test", transform=transforms.ToTensor())
     test_dataloader = DataLoader(
         test_dataset,
-        batch_size=16,
+        batch_size=1,
         num_workers=1,
         shuffle=False,
         pin_memory=(device == "cuda"),
@@ -433,7 +432,7 @@ def main(argv):
         Path(args.output_directory).mkdir(parents=True, exist_ok=True)
 
 
-    model = ELICModel()
+    model = ELICModel(M=192,N=128)
     checkpoint = torch.load(args.checkpoint, map_location=device)
     model.load_state_dict(checkpoint["state_dict"])
     model.update()

### 2023-03-13 model.py

@@ -223,7 +223,7 @@ class ELICModel(JointAutoregressiveHierarchicalPriors):
         indexes = self.gaussian_conditional.build_indexes(scales_squeeze)
         anchor_hat = decoder.decode_stream(indexes.reshape(-1).tolist(), cdf, cdf_lengths, offsets)
         anchor_hat = torch.Tensor(anchor_hat).reshape(scales_squeeze.shape).to(
-            scales_anchor.device) + means_squeeze
+            scales_anchor.device).to(scales_anchor.dtype) + means_squeeze
         anchor_hat = self.ckbd_anchor_unsequeeze(anchor_hat)
         return anchor_hat
 
@@ -239,34 +239,34 @@ class ELICModel(JointAutoregressiveHierarchicalPriors):
         indexes = self.gaussian_conditional.build_indexes(scales_squeeze)
         nonanchor_hat = decoder.decode_stream(indexes.reshape(-1).tolist(), cdf, cdf_lengths, offsets)
         nonanchor_hat = torch.Tensor(nonanchor_hat).reshape(scales_squeeze.shape).to(
-            scales_nonanchor.device) + means_squeeze
+            scales_nonanchor.device).to(scales_nonanchor.dtype) + means_squeeze
         nonanchor_hat = self.ckbd_nonanchor_unsequeeze(nonanchor_hat)
         return nonanchor_hat
 
     def ckbd_anchor_sequeeze(self, y: torch.Tensor) -> torch.Tensor:
         B, C, H, W = y.shape
-        anchor = torch.zeros([B, C, H, W // 2]).to(y.device)
+        anchor = torch.zeros([B, C, H, W // 2], dtype=y.dtype).to(y.device)
         anchor[:, :, 0::2, :] = y[:, :, 0::2, 1::2]
         anchor[:, :, 1::2, :] = y[:, :, 1::2, 0::2]
         return anchor
 
     def ckbd_nonanchor_sequeeze(self, y: torch.Tensor) -> torch.Tensor:
         B, C, H, W = y.shape
-        nonanchor = torch.zeros([B, C, H, W // 2]).to(y.device)
+        nonanchor = torch.zeros([B, C, H, W // 2], dtype=y.dtype).to(y.device)
         nonanchor[:, :, 0::2, :] = y[:, :, 0::2, 0::2]
[…]

### 2023-03-13 modules.py

@@ -78,7 +78,7 @@ class SCCTXModel(nn.Module):
         """
         if block_i == 0:
             bs,c,h,w = psi.shape
-            ch_cx = torch.zeros(bs, 2*self.block_sizes[block_i], h, w)
+            ch_cx = torch.zeros(bs, 2*self.block_sizes[block_i], h, w, dtype=psi.dtype, device=psi.device)
         else:
             ch_cx = self.g_chs[block_i-1](torch.cat(y_hat, dim=1))
 

### 2023-03-13 paper/Template.tex

--- 2023-03-13 paper/paper.tex
@@
 % ------
-\title{Learned image compression}
+\title{AUTHOR GUIDELINES FOR ICIP 2022 PROCEEDINGS MANUSCRIPTS}
 %
@@
 %
-\section{Introduction} % Juha
+\section{Introduction}
 \label{sec:intro}
-With the rise of cloud services and mobile devices, the amount of stored digital media has grown rapidly. 
-Large part of this data is stored images which take considerable amount of space.
-Traditionally these images have been compressed or encoded to formats like jpeg or png which compress the images and usually lose some details (lossy compression).
-With the rise of neural networks and machine learning there has been many attempts to achieve higher compression ratios and lose less details.
-Many of these methods have achieved higher compression ratio with better perceived image quality and less visible artifacts but have been prohibitively expensive in terms of computational cost.
 
-This work tries to replicate paper from Dailan et al\cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
-The model is based on multiple convolutional neural networks which encode the image into latent space.
-This latent representation is then passed into autoencoder which compresses it and is sent to the receiver. 
-The autoencoder takes as parameters input from a model they call Space-Channel ConTeXt (SCCTX).
-This model optimizes the arithmetic encoders performance with input parameters from ... (some technical stuff about scctx) TEEMU?
-
-The output of the arithmetic decoder is passed into otherwise identical neural network as the first one in the pipeline, but in reversed order of layers with transposed convolutional layers instead of convolutional layers.
-The output of this network is now our final image output.
-
-There are still some limitations to this approach. 
-One of which is that the network input image size is limited.
-This means that the network cannot take as input arbitrarily sized images but of course these images can be padded or split or cropped and inputted to the network in multiple pieces.
-Second limitation of this network is the amount of time it takes to encode an image and decode an image. 
-Even when it is considerably faster than other learned image compression methods, in our testing it still very slow and instead of highlighted 100 microseconds for a image which size was not specified, when we tried to encode and compress an 1024x2048 sized image it took approximately 4 seconds for encoding and 5 seconds for decoding.
-Most likely our implementation has many problems which would cause this value to be far away from the highlighted one, but we are about 4 magnitudes away from their highlighted value which at least to us raises some questions.
-
-\section{Experiments}
-\label{sec:experiments}
-\subsection{Training} % Juha
-\label{sec:training}
-
-Training of the model was done with 6000 largest images from ImageNet \cite{imagenet} which were randomly cropped into size of 16x16 pixels.
-These images were then run through the forward method of the network.
-The forward method of the network does not fully compress the images but it transforms them into the latent spaces and the loss function takes into account the likelihood of good compression ratio as well as mse loss.
-
-The model was trained with multiple different configurations.
-We tried $\lambda \in \{0.0008, 0,001\}$ where latent space channels $M$ and $N$ were $M=N=192$ and $\lambda \in \{0.0004, 0.0008\}$ where latent space channels $M$ and $N$ were $M=192, N=128$.
-All of these models were trained for approximately 300 epochs but they could be still trained further.
-Unfortunately due to lack of computational capabilities we could not train more networks and even these networks most likely could have been trained for longer but with epochs taking from 180 seconds (for smaller latent space) to 300 seconds (larger latent space), we could not train them further due to time constraints.
-
-\vfill\pagebreak
-\vfill\pagebreak
+These guidelines include complete descriptions of the fonts, spacing, and
+related information for producing your proceedings manuscripts. Please follow
+them and if you have any questions, direct them to Conference Management
+Services, Inc.: Phone +1-979-846-6800 or email
+to \\\texttt{icip2022@cmsworkshops.com}.
 

### 2023-03-13 paper/paper.tex
Add authors + author's contributions

@@
 % ---------------
-\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
-\address{Author Affiliation(s)}
+\name{Juha Ylikoski, Teemu Eerola, Nardos Estifanos}
+\address{juha.ylikoski@tuni.fi, teemu.eerola@tuni.fi, nardos.estifanos@tuni.fi}
 %
@@
 \begin{keywords}
-One, two, three, four, five
+Learned image compression
 \end{keywords}
 %
+
+\setcounter{section}{-1}
+
+\section{Authors' contributions}
+
 \section{Introduction} % Juha
@@
 
-This work tries to replicate paper from Dailan et al\cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
+This work tries to replicate paper ELIC: Efficient learned image compression with unevenly grouped space-channel contextual adaptive coding \cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
 The model is based on multiple convolutional neural networks which encode the image into latent space.

### 2023-03-14 paper/paper.tex
methods

--- 2023-03-13 paper/paper.tex
@@
 % ---------------
-\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
-\address{Author Affiliation(s)}
+\name{Authors\\ Juha Ylikoski, Nardos Estifanos, Teemu Eerola \\Tampere University\thanks{Thanks to XYZ agency for funding.}}
+\address{\{juha.ylikoski, nardos.estifanos, teemu.eerola\}@tuni.fi}
 %
@@
 \begin{keywords}
-One, two, three, four, five
+AUTHORS CONTRIBUTION?
 \end{keywords}
@@
 
-This work tries to replicate paper from Dailan et al\cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
+This work tries to replicate paper from He \textit{et al.}\cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
 The model is based on multiple convolutional neural networks which encode the image into latent space.
@@
 The autoencoder takes as parameters input from a model they call Space-Channel ConTeXt (SCCTX).
-This model optimizes the arithmetic encoders performance with input parameters from ... (some technical stuff about scctx) TEEMU?
+This model optimizes the encoding/decoding parameters of Gaussian encoder to minimize the redundancy in both channel dimension and in spatial dimension to further reduce the number of bits transferred.
 
@@
 Most likely our implementation has many problems which would cause this value to be far away from the highlighted one, but we are about 4 magnitudes away from their highlighted value which at least to us raises some questions.
+
+\section{Methodology}
+\label{sec:methods}
+\subsection{architecture}
+\label{sec:architecture}
+In this work, we adopted the ELIC-sm model architecture, which is a significantly smaller model that full ELIC architecture presented in Figure \ref{fig:architecture}.
+The differences to ELIC are in main transform networks $g_a$ and $g_s$ where ELIC-sm do not implement attention blocks and have only 1 residual block sequentially connected rather than 3 in full ELIC.
+The architecture of the implemented model is described in table \ref{table:architecture} following table 2 of ELIC paper supplementary material \cite{ELIC}.
+In the table 5x5 denotes the kernel size 5, s2 the stride in convolution to be 2 and the last symbol in each row presents the number of output channels.
+\begin{figure}
+    \centering
+    \includegraphics[width=3in]{architecture.png}
+    \caption{Architecture of full ELIC \cite{ELIC}. The
+blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding. MAKE THE FIGURE AND CAPTION TO SPAN 2 COLUMNS}
+    \label{fig:architecture}
+\end{figure}
+
+\begin{center}
+\caption{Table \ref{table:architecture}: Architecture of ELIC-sm main transform networks. \cite{ELIC}}\\
+\begin{tabular}{l|l}
+\hline
+Analyzer $g_a$ & Synthesizer $g_s$ \\
+\hline
+in: 3-channel image & in: M-channel symbols \\
+\hline
+Conv 5 $\times$ 5, s2, N & TConv 5 $\times$ 5, s2, N \\
+ResBottleneck$\times$1 & ResBottleneck$\times$1 \\
+Conv 5 $\times$ 5, s2, N & TConv 5 $\times$ 5, s2, N \\
+ResBottleneck$\times$1 & ResBottleneck$\times$1 \\
+Conv 5 $\times$ 5, s2, N & TConv 5 $\times$ 5, s2, N \\
+ResBottleneck$\times$1 & ResBottleneck$\times$1 \\
+Conv 5 $\times$ 5, s2, M & TConv 5 $\times$ 5, s2, 3 \\
+\hline
+\end{tabular}
+\label{table:architecture}
+\end{center}
+
+We chose the smaller ELIC-sm architecture to reduce the amount of computational complexity and the need for data.
+We also kept the hyperparameters $M$ and $N$ reasonably low to further ease the training of the model.
+$M$ controls the number of channels in latent space and $N$ the number of channels in main transform networks.
+We leverage the ready-made implementation of Gaussian and arithmetic en/de-coders from CompressAI library \cite{compressai}.
+
+\subsection{Space-Channel Context model}
+In works previously that ELIC \cite{ELIC} the context models were used only to predict either parameters of channel wise or spatial context to encode/decode pixels \cite{balle2016, mbt2018, balle2018, checkerboard}.
+The proposed method in ELIC uses both channel and spatial information in context model to predict the parameters $\mathbf{\Theta_i^{(k)}}$of next group \cite{ELIC}.
+
+The channel groups are formed in uneven manner based on observation that most of the energy is often concentrated on early channels.
+Therefore, the number of channels in groups is rising towards latter groups.
+The context prediction is made using all previous channel groups combined.
+When encoding or decoding the first channel group the channel wise context is therefore set to zeros.
+Because the uneven groups the sizes of inputs and outputs of the channel context model changes over the groups, and therefore separate model is implemented for each group.
+The number of networks is one less than number of groups because the networks make prediction of the next group parameters.
+The architecture of channel-wise context model is implemented as it is from the ELIC paper.
+\cite{ELIC} We doubled the number of channels only in the last convolutional layer.
+
+The spatial context can be implemented as autoregressive convolution where each pixel prior to the current pixel must be predicted \cite{mbt2018}.
+A more efficient approach to implement spatial context prediction as checkerboard convolution introduced by He \textit{et al.} 2021 \cite{checkerboard}.
+The concept involves dividing the group into half like picking only the white or black squares from checkerboard and making predictions to these anchor pixels first.
+Another half of the group, the non-anchors, can be then predicted using the firstly encoded/decoded group as context.
+In the prediction of anchor parameters, the spatial context is simply set to zeros.
+The architecture of spatial context module is implemented with one checkerboard convolution layer with kernel size of 5 and stride of 1 that keeps the number of channels the same.
+
+The combined representation of prediction parameters is got from parameter aggregation network that combines the spatial and channel wise context predictions with hyperprior $\mathbf{\Psi}$ from hyperparameter synthetization network $h_s$.
+Because of uneven sizes of channel groups, channel and spatial context prediction networks, a separate networks needs to be implemented for each group.
+Architecture of parameter aggregation network is implemented as it is in the ELIC paper figure 6 with linear reduction of channels in layers \cite{ELIC}.
+
+The table \ref{table:SCCTX} shows the number of trainable parameters in thousands of SCCTX for channel number of $M=192$ and the default channel block sizes $(16,16,32,64,M-128)$ from the ELIC paper.
+The channel context model parameters for block$_0$ is \textit{NaN}, because it does not exists due to reasons discussed earlier.
+
+\begin{center}
+\caption{Table \ref{table:SCCTX}: Number of trainable parameters in SCCTX in\\ thousands for each block ($b_i$) and for each component.}\\
+\begin{tabular}{c c c c c c c} 
+ \hline
+ _ & b_0 & b_1 & b_2 & b_3 & b_4 & sum \\  
+ \hline\hline
+ channel & \textit{NaN} & 26 & 103 & 410 & 512 & 1050 \\
+ \hline
+ spatial & 13 & 13 & 51 & 205 & 205 & 487 \\
+ \hline
+ par.agg. & 198 & 198 & 277 & 480 & 480 & 1632 \\
+ \hline
+ sum & 210 & 236 & 431 & 1095 & 1197 & 3169 \\
+ \hline
+\end{tabular}
+\label{table:SCCTX}
+\end{center}
+
+\subsection{Training versus compression flow}
+The inference and training procedures differ from each other due to the undifferentiable nature of encoders.
+In inference time we want to compress the image to actual discrete bitstream and in training we want to just get the likelihoods from the entropy models and retain the gradient information.
+The gradient can be retained from the encoders by not compressing the image to bitstream, but just adding gaussian noise to image to mimic the effect of quantization.
+The approach was presented by \cite{balle2016}.
+With the checkerboard spatial context model another shortcut can be made in training time to further reduce the computational complexity of training.
+Following the work of He \textit{et al.} 2021 \cite{checkerboard} we used one pass encoding in training where parameters of non-anchor blocks are calculated from the input without calculating the parameters of anchor blocks.
+This approach was different from that presented in ELIC paper \cite{ELIC}, where they encoded $y+mu$ to bitstream instead of $y$ and therefore one pass encoding was not possible.
+We did not follow the implementation of ELIC because of lack of understanding about entropy models.
+
+At inference time we calculate a separate bitstream to each image in batch sequentially and leverage the full context prediction by first decoding the anchor blocks and then the non-anchor blocks.
+This still allows us to decode the channel blocks in the image separately using the features of entropy model.
+However, for practical use cases, the batch processing should be allowed to benefit from the parallel computation capabilities of GPUs.
 

### 2023-03-14 paper/paper.tex

@@
 % ---------------
-\name{Authors\\ Juha Ylikoski, Nardos Estifanos, Teemu Eerola \\Tampere University\thanks{Thanks to XYZ agency for funding.}}
-\address{\{juha.ylikoski, nardos.estifanos, teemu.eerola\}@tuni.fi}
+\name{Juha Ylikoski, Teemu Eerola, Nardos Estifanos}
+\address{juha.ylikoski@tuni.fi, teemu.eerola@tuni.fi, nardos.estifanos@tuni.fi}
 %
@@
 \begin{keywords}
-AUTHORS CONTRIBUTION?
+Learned image compression
 \end{keywords}
 %
+
+\setcounter{section}{-1}
+
+\section{Authors' contributions}
+
 \section{Introduction} % Juha
@@
 
-This work tries to replicate paper from He \textit{et al.}\cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
+This work tries to replicate paper ELIC: Efficient learned image compression with unevenly grouped space-channel contextual adaptive coding \cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
 The model is based on multiple convolutional neural networks which encode the image into latent space.

### 2023-03-14 paper/paper.tex
conclusions

@@
 The model was trained with multiple different configurations.
-We tried $\lambda \in \{0.0008, 0,001\}$ where latent space channels $M$ and $N$ were $M=N=192$ and $\lambda \in \{0.0004, 0.0008\}$ where latent space channels $M$ and $N$ were $M=192, N=128$.
+We tried $\lambda \in \{0.0008, 0.001\}$ where latent space channels $M$ and $N$ were $M=N=192$ and $\lambda \in \{0.0004, 0.0008\}$ where latent space channels $M$ and $N$ were $M=192, N=128$.
 All of these models were trained for approximately 300 epochs but they could be still trained further.
 Unfortunately due to lack of computational capabilities we could not train more networks and even these networks most likely could have been trained for longer but with epochs taking from 180 seconds (for smaller latent space) to 300 seconds (larger latent space), we could not train them further due to time constraints.
+
+\section{Discussions and Conclusions}
+The goal of this work was to implement and verify the methods of He \textit{et al.} 2022 \cite{ELIC} for learned image compression.
+We successfully implemented and trained 4 different models targeted to low compression ratios by varying the model hyperparameters $M$, $N$ and $\lambda$ in loss function.
+The implementation differs from the paper in few details as discussed earlier.
+Due to limited time and computing resources we were not able to train our models to convergence and we used fewer images in our training set.
+Finetuning of the models using MS-SSIM as loss function was also discarded due to computation and time restrictions.
+
+The results from our models were far from the level of original ELIC paper in terms of subjective and objective measurements in quality.
+This was caused by implementation differences and different training setups.
+Additionally, our models suffer from color artifacts that were introduced in various areas in result images.
+We are not sure what is the cause of these artifacts, but possible reasons are errors in implementation for example overflow in data types or underfitted models.
+In subjective evaluation we were able to pass the level of JPEG compression, but objective metrics were orders of magnitude behind JPEG.
+We can conclude that were where able to have working image compression models using methods introduced by He \textit{et al.} 2022 \cite{ELIC} in ELIC, but we can not confirm the improved quality and inference speed they achieved.
+We neither studied the progressive decoding or the fast thumbnail generator introduced in original paper.
+
+Apparently the models from original ELIC paper also suffers from some kind of artifacts that they do not mention in paper.
+In successive work He \textit{et al.} 2022 \cite{PO-ELIC} propose a discriminator network in top of ELIC to reduce amount of know artifacts in learned image compression with Generative adversarial network (GAN) inspired training.
+Other interesting directions in learned image compression are purely GAN and diffusion based compressors, that are capable of reconstructing even unseen areas of image using techniques of generative methods \cite{HIFIC, diffusion}.
 

### 2023-03-14 paper/paper.tex
Fix tables

@@
 
+\begin{table}[!h]
+\label{table:architecture}
 \begin{center}
-\caption{Table \ref{table:architecture}: Architecture of ELIC-sm main transform networks. \cite{ELIC}}\\
+\caption{Architecture of ELIC-sm main transform networks. \cite{ELIC}}
 \begin{tabular}{l|l}
@@
 \end{tabular}
-\label{table:architecture}
 \end{center}
+\end{table}
 
@@
 
+\begin{table}[!h]
+\label{table:SCCTX}
 \begin{center}
-\caption{Table \ref{table:SCCTX}: Number of trainable parameters in SCCTX in\\ thousands for each block ($b_i$) and for each component.}\\
+\caption{Number of trainable parameters in SCCTX in\\ thousands for each block ($b_i$) and for each component.}
 \begin{tabular}{c c c c c c c} 
  \hline
- _ & b_0 & b_1 & b_2 & b_3 & b_4 & sum \\  
+  & $b_0$ & $b_1$ & $b_2$ & $b_3$ & $b_4$ & sum \\
  \hline\hline
@@
 \end{tabular}
-\label{table:SCCTX}
 \end{center}
+\end{table}
 

### 2023-03-14 paper/paper.tex
Add abstract, sort + format references

@@
 \begin{abstract}
-The abstract should appear at the top of the left-hand column of text, about
-0.5 inch (12 mm) below the title area and no more than 3.125 inches (80 mm) in
-length.  Leave a 0.5 inch (12 mm) space between the end of the abstract and the
-beginning of the main text.  The abstract should contain about 100 to 150
-words, and should be identical to the abstract text submitted electronically
-along with the paper cover sheet.  All manuscripts must be in English, printed
-in black ink.
+Image compression has had some major new ideas in forms of learned image compression. 
+Especially the neural network based learned image compression methods have achieved even better performance than the more traditional manually designed lossy image compression models.
+These new methods are not without their own problems.
+The large neural network based models can achieve remarkable compression ratios and perceived image quality, but have suffered widely form being too computationally expensive to be usable.
+
+In this work we are reviewing and trying to replicate one of these methods which claims good quality, compression ratio and speed.
 \end{abstract}
@@
 \begin{keywords}
-Learned image compression
+Learned image compression, ELIC Efficient Learned Image Compression
 \end{keywords}

### 2023-03-15 eval_mode.py
* modified evaluation (now takes all models at once)

@@ -1,17 +1,11 @@
-# python eval_mode.py checkpoint  --checkpoint "./checkpoint_best_loss.pth.tar" --dataset "./CLIC Dataset/" -d "./output_eval_clic"
-# python eval_mode.py checkpoint  --checkpoint "./checkpoint_best_loss.pth.tar" --dataset "./KODAK Dataset/" -d "./output_eval_kodak"
-
-# python plot.py -f './output_eval_kodak/ELIC MODEL.json' -t "ELIC Model Evaluation on KODAK dataset [PSNR]" -o './output_eval_kodak/elic_kodak_psnr.png' --show
-# python plot.py -f './output_eval_kodak/ELIC MODEL.json' -t "ELIC Model Evaluation on KODAK dataset [MS-SSIM]" -o './output_eval_kodak/elic_kodak_mssim.png' -m 'ms-ssim' --show
-
-# python plot.py -f './output_eval_kodak/ELIC MODEL.json' -t "ELIC Model Evaluation on CLIC dataset [PSNR]" -o './output_eval_kodak/elic_clic_psnr.png' --show
-# python plot.py -f './output_eval_kodak/ELIC MODEL.json' -t "ELIC Model Evaluation on CLIC dataset [MS-SSIM]" -o './output_eval_kodak/elic_clic_msssim.png' -m 'ms-ssim' --show
+# python eval_mode.py checkpoint  --dataset "./CLIC Dataset/" -d "./output_eval_clik" -of "elic_clic"  --checkpoints "./[250,8e-4]_checkpoint_best_loss.pth.tar" "./[300,4e-4]_checkpoint_best_loss.pth.tar" "./[300,8e-4]_checkpoint_best_loss.pth.tar" "./[300,good]checkpoint_best_loss.pth.tar"  --N 192 128 128 192 --M 192 192 192 192
 
 import argparse
 import json
 import math
 import sys
 import time
+import os
 from collections import defaultdict
 from pathlib import Path
 from typing import Any, Dict, List
@@ -124,7 +118,7 @@ def inference(model, test_dataloader, criterion, device, loss, bpp_loss, mse_los
         tqdm.write(f"Processing batches {offset // test.shape[0]}")
 
         for i in tqdm(range(test.shape[0])):
-
+            
             x = test[i]
             x = x.unsqueeze(0)
 
@@ -155,20 +149,21 @@ def inference(model, test_dataloader, criterion, device, loss, bpp_loss, mse_los
             
[…]

### 2023-03-15 paper/paper.tex
* related works & training last part added

@@
 Most likely our implementation has many problems which would cause this value to be far away from the highlighted one, but we are about 4 magnitudes away from their highlighted value which at least to us raises some questions.
+
+\section{Related Works}
+\label{sec:related_works}
+Recently, in learned compression, we have seen the proliferation of deep learning-based image compression techniques using automatic coding architectures with great success and promising results.
+The development of previous learning-based compression models has spanned several years, hence, there are many related works.
+
+In the early stages, some work was proposed to perform non-differential quantization and rate estimation for complete training, for instance, some works focused on developing network structures that can obtain more compact and efficient latent images and reconstruct high-quality images from compressed images.
+
+It is noted that traditional image compression techniques, such as JPEG and BPG, have limitations in terms of subjective quality and rate-distortion performance.
+Learned image compression techniques have shown promise in addressing these limitations by approaching it as rate-distortion optimization problem \cite{ELIC}.
+
+\begin{equation}
+    R + \lambda \cdot D
+\end{equation}
+
+Some approaches include content weighting strategies or combining of different channels using principal component analysis, or consider energy compression or use deep residual devices to improve the network architecture \cite{balle2016}.
+Recently, several studies have investigated adaptive contextual entropy estimation models to guide the optimization process of neural network parameters to achieve an optimal trade-off between reconstruction error and required bits (entropy), including \cite{AgustssonEirikur2017SVQf, balle2018, mbt2018, LeeJooyoung2019CEMf} Entropy estimation methods have greatly improved learning compression algorithms, and the most representative methods are hyperprior models and joint models.
+
+Several recent works that have made significant contributions to the field of learned image compression.
+One such work is the \cite{VincentPascal2008Eacr, TheisLucas2017Licw, balle2016} End-to-End Optimized Image Compression (EOIC) model, which uses a convolution neural network (CNN) to compress images.
+EOIC uses a combination of entropy coding and arithmetic coding to achieve high compression rates.
+Another work is the Generative Adversarial Networks (GAN)-based image compression model, which uses a GAN to generate compressed images.
+GAN-based models have shown promise in improving the subjective quality of compressed images.
+
+The use of attention mechanisms in learned image compression models is also another discussion.
+Attention mechanisms allow the model to focus on specific regions of the image that are important for compression.
+One such model is the Attention Recurrent Network for Image Compression (ARNIC) \cite{XueYuyang2019ABIC}, which uses a recurrent neural network (RNN) with an attention mechanism to compress images.
+Attention mechanisms have shown promise in improving the rate-distortion performance of learned image compression models.
+
+Another important improvement comes from the use of residual blocks in learned image compression models \cite{balle2016}.
+Residual blocks are a type of neural network layer that allows for the reuse of features learned in previous layers.
+They have shown promise in improving the rate-distortion performance of learned image compression models.
+They are also easier to extend for dynamic or slimmable inference, while other techniques like GDN require special handling.
+
+Learning from these techniques have shown promise in improving the rate-distortion performance of learned image compression models.
+It is proposed to make use of uneven channel-conditional adaptive coding, motivated by the observation of energy compaction in learned image compression.
 
@@
 Unfortunately due to lack of computational capabilities we could not train more networks and even these networks most likely could have been trained for longer but with epochs taking from 180 seconds (for smaller latent space) to 300 seconds (larger latent space), we could not train them further due to time constraints.
+
+After obtaining different models with different parameters, we have calculated the PSNR and MS-SSIM \cite{Bjntegaard2001CalculationOA, WangZhou2003Mssf} values to plot the rate-distortion curve which helps in visualizing the performance of compression as well as perceived quality and bits used to represent an image compared to each other and other techniques such as JPEG.
+For this evaluation, we have modified the script from compressai \cite{compressai} to get close enough comparison with other techniques, and made use of KODAK\cite{kodak} and CLIC professional \cite{clic} datasets.
+
+\section{Results}
+\label{sec:results}
+
+\begin{figure}[htb]
+
+\begin{minipage}[b]{1.0\linewidth}
+  \centering
+  \centerline{\includegraphics[width=8.5cm]{image1}}
+%  \vspace{2.0cm}
+  \centerline{(a) Result 1}\medskip
+\end{minipage}
+%
+\begin{minipage}[b]{.48\linewidth}
+  \centering
+  \centerline{\includegraphics[width=4.0cm]{image3}}
+%  \vspace{1.5cm}
+  \centerline{(b) Results 3}\medskip
+\end{minipage}
+\hfill
+\begin{minipage}[b]{0.48\linewidth}
+  \centering
+  \centerline{\includegraphics[width=4.0cm]{image4}}
+%  \vspace{1.5cm}
+  \centerline{(c) Result 4}\medskip
+\end{minipage}
+%
+\caption{Example of placing a figure with experimental results.}
+\label{fig:res}
+%
+\end{figure}
 

### 2023-03-15 paper/paper.tex
added BD-section

@@
 
+\subsection{Evaluation}
 After obtaining different models with different parameters, we have calculated the PSNR and MS-SSIM \cite{Bjntegaard2001CalculationOA, WangZhou2003Mssf} values to plot the rate-distortion curve which helps in visualizing the performance of compression as well as perceived quality and bits used to represent an image compared to each other and other techniques such as JPEG.
@@
 \end{figure}
+
+To compare the rate distortion performance of various models, we
+present BD-rate \cite{Bjntegaard2001CalculationOA} computed from PSNR-BPP curves as
+the quantitative metric. 
+Like seen from the Fig.
+\ref{fig:res} our models performance in quantitative evaluation is much worse than any conventional method.
+The BD-SNR is 19.1929 dB and BD-rate -3.367e+84\% compared to JPEG.
+In calculations we used only JPEG compressors those \textit{bdd} were in the same range than your models.
+
+From our models, we can compare the performance of two models that were trained on $\lambda$=8e-4 by using the KODAK dataset \cite{kodak}.
+Other models are not comparable in with BD-rates because the range of \textit{bbp} are different.
+In terms of BD-rate we got 3.987\% from the model that was trained only 250 epochs compared to the one trained for 300 epochs with yields that less trained model actually performs better.
+The reason for this is unknown, but overfitting might not an option because the models were still unconverged.
 

### 2023-03-15 paper/paper.tex
* results & images added
Reorganize images

--- 2023-03-15 paper/paper.tex
@@
 \usepackage{spconf,amsmath,graphicx}
+
+\usepackage[section]{placeins}
+\usepackage{float}
 
@@
 In the table 5x5 denotes the kernel size 5, s2 the stride in convolution to be 2 and the last symbol in each row presents the number of output channels.
-\begin{figure}
+
+\begin{figure*}[h!]
     \centering
-    \includegraphics[width=3in]{architecture.png}
+    \includegraphics[width=7in]{architecture.png}
     \caption{Architecture of full ELIC \cite{ELIC}. The
@@
     \label{fig:architecture}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[htb]
+In Figure a, we demonstrate the reconstruction performance of the four models trained with different parameters.
 
-\begin{minipage}[b]{1.0\linewidth}
+\pagebreak
+
+% \begin{minipage}
+% \begin{flushleft}
+% \end{flushleft}
+% \begin{itemize}
+%     \item[] Row (Model) 1: epoch=250 | $\lambda$=8e-4 | N=M=192, 
+%     \item[] Row (Model) 2: epoch=300 | $\lambda$=4e-4 | N=128, M=192, 
+%       \item[] Row (Model) 3: epoch=300 | $\lambda$=8e-4 | N=128, M=192, 
+%       \item[] \textbf{Row (Model) 4}: epoch=300 | $\lambda$=1e-2 | N=M=192
+% \end{itemize}
+% \end{flushleft}
+% \end{minipage}
+
+\begin{figure}[h!]
+\begin{minipage}[b]{0.8\linewidth}
     \centering
-  \centerline{\includegraphics[width=8.5cm]{image1}}
-%  \vspace{2.0cm}
-  \centerline{(a) Result 1}\medskip
+    \includegraphics[width=7.5cm]{psnr_mssim.png}
 \end{minipage}
-%
-\begin{minipage}[b]{.48\linewidth}
+\begin{minipage}[b]{0.8\linewidth}
+
     \centering
-  \centerline{\includegraphics[width=4.0cm]{image3}}
-%  \vspace{1.5cm}
-  \centerline{(b) Results 3}\medskip
+    \includegraphics[width=7.5cm]{per_img.png}
 \end{minipage}
-\hfill
-\begin{minipage}[b]{0.48\linewidth}
+    \label{fig:architecture}
+    \caption{Architecture of full ELIC \cite{ELIC}. The
+blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding. MAKE THE FIGURE AND CAPTION TO SPAN 2 COLUMNS}
+\end{figure}
+
+\pagebreak
+
+\begin{figure*}[h!]
 \centering
-  \centerline{\includegraphics[width=4.0cm]{image4}}
-%  \vspace{1.5cm}
-  \centerline{(c) Result 4}\medskip
+\begin{minipage}[b]{.8\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{reconstruction.png}
+
 \end{minipage}
-%
-\caption{Example of placing a figure with experimental results.}
-\label{fig:res}
-%
-\end{figure}
+\begin{minipage}[b]{.5\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{comparison.png}
+\end{minipage}
+    \caption{Architecture of full ELIC \cite{ELIC}. The
+blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding. MAKE THE FIGURE AND CAPTION TO SPAN 2 COLUMNS}
+    \label{fig:architecture}
+\end{figure*}
+
+\vfill\pagebreak
+\vfill\pagebreak
 
@@
 
-\vfill\pagebreak
-\vfill\pagebreak
-
-\section{Formatting your paper}
-\label{sec:format}
-
-All printed material, including text, illustrations, and charts, must be kept
-within a print area of 7 inches (178 mm) wide by 9 inches (229 mm) high. Do
-not write or print anything outside the print area. The top margin must be 1
-inch (25 mm), except for the title page, and the left margin must be 0.75 inch
-(19 mm).  All {\it text} must be in a two-column format. Columns are to be 3.39
-inches (86 mm) wide, with a 0.24 inch (6 mm) space between them. Text must be
-fully justified.
-
-\section{PAGE TITLE SECTION}
-\label{sec:pagestyle}
-
-The paper title (on the first page) should begin 1.38 inches (35 mm) from the
-top edge of the page, centered, completely capitalized, and in Times 14-point,
-boldface type.  The authors' name(s) and affiliation(s) appear below the title
-in capital and lower case letters.  Papers with multiple authors and
-affiliations may require two or more lines for this information. Please note
-that papers should not be submitted blind; include the authors' names on the
-PDF.
-
-\section{TYPE-STYLE AND FONTS}
-\label{sec:typestyle}
-
-To achieve the best rendering both in printed proceedings and electronic proceedings, we
-strongly encourage you to use Times-Roman font.  In addition, this will give
-the proceedings a more uniform look.  Use a font that is no smaller than nine
-point type throughout the paper, including figure captions.
-
-In nine point type font, capital letters are 2 mm high.  {\bf If you use the
-smallest point size, there should be no more than 3.2 lines/cm (8 lines/inch)
-vertically.}  This is a minimum spacing; 2.75 lines/cm (7 lines/inch) will make
-the paper much more readable.  Larger type sizes require correspondingly larger
-vertical spacing.  Please do not double-space your paper.  TrueType or
-Postscript Type 1 fonts are preferred.
-
-The first paragraph in each section should not be indented, but all the
-following paragraphs within the section should be indented as these paragraphs
-demonstrate.
-
-\section{MAJOR HEADINGS}
-\label{sec:majhead}
-
-Major headings, for example, "1. Introduction", should appear in all capital
-letters, bold face if possible, centered in the column, with one blank line
-before, and one blank line after. Use a period (".") after the heading number,
-not a colon.
-
-\subsection{Subheadings}
-\label{ssec:subhead}
-
-Subheadings should appear in lower case (initial word capitalized) in
-boldface.  They should start at the left margin on a separate line.
-
-\subsubsection{Sub-subheadings}
-\label{sssec:subsubhead}
-
-Sub-subheadings, as in this paragraph, are discouraged. However, if you
-must use them, they should appear in lower case (initial word
-capitalized) and start at the left margin on a separate line, with paragraph
-text beginning on the following line.  They should be in italics.
-
-\section{PRINTING YOUR PAPER}
-\label{sec:print}
-
-Print your properly formatted text on high-quality, 8.5 x 11-inch white printer
-paper. A4 paper is also acceptable, but please leave the extra 0.5 inch (12 mm)
-empty at the BOTTOM of the page and follow the top and left margins as
-specified.  If the last page of your paper is only partially filled, arrange
-the columns so that they are evenly balanced if possible, rather than having
-one long column.
-
-In LaTeX, to start a new column (but not a new page) and help balance the
-last-page column lengths, you can use the command ``$\backslash$pagebreak'' as
-demonstrated on this page (see the LaTeX source below).
-
-\section{PAGE NUMBERING}
-\label{sec:page}
-
-Please do {\bf not} paginate your paper.  Page numbers, session numbers, and
-conference identification will be inserted when the paper is included in the
-proceedings.
-
-\section{ILLUSTRATIONS, GRAPHS, AND PHOTOGRAPHS}
-\label{sec:illust}
-
-Illustrations must appear within the designated margins.  They may span the two
-columns.  If possible, position illustrations at the top of columns, rather
-than in the middle or at the bottom.  Caption and number every illustration.
-All halftone illustrations must be clear black and white prints.  Colors may be
-used, but they should be selected so as to be readable when printed on a
-black-only printer.
-
-Since there are many ways, often incompatible, of including images (e.g., with
-experimental results) in a LaTeX document, below is an example of how to do
-this \cite{Lamp86}.
-
-\section{FOOTNOTES}
-\label{sec:foot}
-
-Use footnotes sparingly (or not at all!) and place them at the bottom of the
-column on the page on which they are referenced. Use Times 9-point type,
-single-spaced. To help your readers, avoid using footnotes altogether and
-include necessary peripheral observations in the text (within parentheses, if
-you prefer, as in this sentence).
-
-% Below is an example of how to insert images. Delete the ``\vspace'' line,
-% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
-% with a suitable PostScript file name.
-% -------------------------------------------------------------------------
-\begin{figure}[htb]
-
-\begin{minipage}[b]{1.0\linewidth}
-  \centering
-  \centerline{\includegraphics[width=8.5cm]{image1}}
-%  \vspace{2.0cm}
-  \centerline{(a) Result 1}\medskip
-\end{minipage}
-%
-\begin{minipage}[b]{.48\linewidth}
-  \centering
-  \centerline{\includegraphics[width=4.0cm]{image3}}
-%  \vspace{1.5cm}
-  \centerline{(b) Results 3}\medskip
-\end{minipage}
-\hfill
-\begin{minipage}[b]{0.48\linewidth}
-  \centering
-  \centerline{\includegraphics[width=4.0cm]{image4}}
-%  \vspace{1.5cm}
-  \centerline{(c) Result 4}\medskip
-\end{minipage}
-%
-\caption{Example of placing a figure with experimental results.}
-\label{fig:res}
-%
-\end{figure}
-
-% To start a new column (but not a new page) and help balance the last-page
-% column length use \vfill\pagebreak.
-% -------------------------------------------------------------------------
-%\vfill
-%\pagebreak
-
-\section{COPYRIGHT FORMS}
-\label{sec:copyright}
-
-You must submit your fully completed, signed IEEE electronic copyright release
-form when you submit your paper. We {\bf must} have this form before your paper
-can be published in the proceedings.
-
-\section{RELATION TO PRIOR WORK}
-\label{sec:prior}
-
-The text of the paper should contain discussions on how the paper's
-contributions are related to prior work in the field. It is important
-to put new work in  context, to give credit to foundational work, and
-to provide details associated with the previous work that have appeared
-in the literature. This discussion may be a separate, numbered section
-or it may appear elsewhere in the body of the manuscript, but it must
-be present.
-
-You should differentiate what is new and how your work expands on
-or takes a different path from the prior studies. An example might
-read something to the effect: "The work presented here has focused
-on the formulation of the ABC algorithm, which takes advantage of
-non-uniform time-frequency domain analysis of data. The work by
-Smith and Cohen \cite{Lamp86} considers only fixed time-domain analysis and
-the work by Jones et al \cite{C2} takes a different approach based on
-fixed frequency partitioning. While the present study is related
-to recent approaches in time-frequency analysis [3-5], it capitalizes
-on a new feature space, which was not considered in these earlier
-studies."
-
-\vfill\pagebreak
-
-\section{REFERENCES}
-\label{sec:refs}
-
-List and number all bibliographical references at the end of the
-paper. The references can be numbered in alphabetic order or in
-order of appearance in the document. When referring to them in
-the text, type the corresponding reference number in square
-brackets as shown at the end of this sentence \cite{C2}. An
-additional final page (the fifth page, in most cases) is
-allowed, but must contain only references to the prior
-literature.
-
 % References should be produced using the bibtex program from suitable

### 2023-03-15 jpeg_model.py
Add jpeg + own model evaluation stuff

@@ -1,0 +1,31 @@
+
+import numpy as np
+import imageio.v3 as iio
+from simplejpeg import decode_jpeg, encode_jpeg
+
+
+
+
+def make_bad(path, quality):
+    im = iio.imread(path)
+    compressed = encode_jpeg(im, quality=quality)
+    print(type(compressed))
+    bad = decode_jpeg(compressed)
+    return bad, compressed
+
+
+def psnr(a: np.array, b: np.array, max_val: int = 255) -> float:
+    return 20 * np.log10(max_val) - 10 * np.log10(np.power((a - b), 2).mean())
+
+
+bad, compressed = make_bad("/tmp/test/test.png", 0)
+orig = iio.imread("/tmp/test/test.png")
+iio.imwrite("bad.jpg", bad)
+print(psnr(orig, bad))
+
+num_pixels = 512*512
+
+bpp = len(compressed) * 8.0 / num_pixels
+
+print(bpp)
+

### 2023-03-19 eval_mode copy.py
• evaluation bug fixed • new image added

@@ -47,14 +47,10 @@ def custom_collate(batch):
         img = batch[i] # (3, h, w)
 
         # get the closest power of 2 for the height and width
-        # h = int(2**np.floor(np.log2(img.shape[1])))
-        # w = int(2**np.floor(np.log2(img.shape[2])))
+        h = int(2**np.floor(np.log2(img.shape[1])))
+        w = int(2**np.floor(np.log2(img.shape[2])))
         
-
-
-        
-        # max_dim = max(h,w, max_dim)
-        max_dim = 512
+        max_dim = max(h,w, max_dim)
         
         # resize the image keeping the aspect ratio with skimage
         img = resize(img.numpy(), (3, max_dim, max_dim), preserve_range=True, anti_aliasing=True)
@@ -127,33 +123,33 @@ def inference(model, test_dataloader, criterion, device, loss, bpp_loss, mse_los
             x = x.unsqueeze(0)
 
             h, w = x.size(2), x.size(3)
-            # pad, unpad = compute_padding(h, w, min_div=2**6)  # pad to allow 6 strides of 2
+            pad, unpad = compute_padding(h, w, min_div=2**6)  # pad to allow 6 strides of 2
             
-            # x_padded = F.pad(x, pad, mode="constant", value=0)
+            x_padded = F.pad(x, pad, mode="constant", value=0)
             
             x = x.to(device)
             out_net = model(x)
 
             start = time.time()
-            out_enc = model.compress(x)
+            out_enc = model.compress(x_padded)
             enc_time = time.time() - start
             
             start = time.time()
             out_dec = model.decompress(out_enc['strings'],out_enc['shape'])
             dec_time = time.time() - start
             
-            # out_dec["x_hat"] = F.pad(out_dec["x_hat"], unpad)
+            out_dec["x_hat"] = F.pad(out_dec["x_hat"], unpad)
 
             # input images are 8bit RGB for now
             _metrics = compute_metrics(x, out_dec["x_hat"], 255)
-            print(f'x.size: {x.size()}')
-            num_pixels = x.size(2) * x.size(3)
+            num_pixels = x.size(0) * x.size(2) * x.size(3)
             bpp = sum(len(s[0]) for s in out_enc["strings"]) * 8.0 / num_pixels
             
             print(f'bpp: {bpp}')
             
[…]

### 2023-03-20 paper/paper.tex
update

@@
     \caption{Architecture of full ELIC \cite{ELIC}. The
-blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding. MAKE THE FIGURE AND CAPTION TO SPAN 2 COLUMNS}
+blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding.}
     \label{fig:architecture}
@@
 
-\pagebreak
+\begin{table}[!h]
+\label{table:models}
+\begin{center}
+\begin{tabular}{c c c}
+    Row (Model) 1: epoch=250 & $\lambda$=8e-4 & N=M=192 \\
+    Row (Model) 2: epoch=300 & $\lambda$=4e-4 & N=128, M=192 \\
+    Row (Model) 3: epoch=300 & $\lambda$=8e-4 & N=128, M=192 \\
+    \textbf{Row (Model) 4}: epoch=300 & $\lambda$=1e-2 & N=M=192 \\
+\end{tabular}
+\end{center}
+\end{table}
 
-% \begin{minipage}
-% \begin{flushleft}
-% \end{flushleft}
-% \begin{itemize}
-%     \item[] Row (Model) 1: epoch=250 | $\lambda$=8e-4 | N=M=192, 
-%     \item[] Row (Model) 2: epoch=300 | $\lambda$=4e-4 | N=128, M=192, 
-%       \item[] Row (Model) 3: epoch=300 | $\lambda$=8e-4 | N=128, M=192, 
-%       \item[] \textbf{Row (Model) 4}: epoch=300 | $\lambda$=1e-2 | N=M=192
-% \end{itemize}
-% \end{flushleft}
-% \end{minipage}
+\vfill
 
@@
     \centering
-    \includegraphics[width=7.5cm]{psnr_mssim.png}
+    \includegraphics[width=7.2cm]{kodak_fixed.png}
+    \centerline{(a) Average values over KODAK dataset \cite{kodak}.}\medskip
 \end{minipage}
@@
     \centering
-    \includegraphics[width=7.5cm]{per_img.png}
+    \includegraphics[width=7.2cm]{kodak_per_fixed.png}
+    \centerline{(b) Performance on individual images of KODAK dataset \cite{kodak}.}\medskip
 \end{minipage}
-    \label{fig:architecture}
-    \caption{Architecture of full ELIC \cite{ELIC}. The
-blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding. MAKE THE FIGURE AND CAPTION TO SPAN 2 COLUMNS}
+    \label{fig:results}
+    \caption{Results on KODAK dataset \cite{kodak}.}
 \end{figure}
@@
     \includegraphics[width=\linewidth]{reconstruction.png}
+    \centerline{(a)  Result images from our models. Each row is results from the model 1 to 4 respectively}\medskip
 
@@
     \includegraphics[width=\linewidth]{comparison.png}
+    \centerline{(b) Comparisons to conventional image compression methods JPEG and BPG to our results from Model 3 and original image.}\medskip
 \end{minipage}
-    \caption{Architecture of full ELIC \cite{ELIC}. The
-blue and red arrows denote the encoding and decoding data flow. The orange ones are shared by both encoding and decoding. MAKE THE FIGURE AND CAPTION TO SPAN 2 COLUMNS}
-    \label{fig:architecture}
+    \caption{Result/comparison images.}
+    \label{fig:images}
 \end{figure*}

### 2023-03-20 paper/paper.tex
Fig figures.
* added a paragraph in results section

@@
 
-In Figure a, we demonstrate the reconstruction performance of the four models trained with different parameters.
+In Fig 2, we demonstrate the reconstruction performance of the four models trained with different parameters.
+We can see we're neck-in-neck with the conventional approaches for low bpp rates.
+And Fig 3a.
+we observed Model 4, has better sharpness but suffers from glitch effect in uniform backgrounds.
 
@@
 
-\vfill
+\pagebreak
 
@@
     \includegraphics[width=7.2cm]{kodak_fixed.png}
-    \centerline{(a) Average values over KODAK dataset \cite{kodak}.}\medskip
+    \centerline{(a) Average values.}\medskip
 \end{minipage}
@@
     \includegraphics[width=7.2cm]{kodak_per_fixed.png}
-    \centerline{(b) Performance on individual images of KODAK dataset \cite{kodak}.}\medskip
+    \centerline{(b) Performance on individual images.}\medskip
 \end{minipage}

### 2023-03-20 paper/paper.tex
Add Teemus results + authors contributions

@@
 \section{Authors' contributions}
+
+Juha Ylikoski was responsible for creating the cnns around the SCCTX-model and training the networks.
+Teemu Eerola was responsible for creating the SCCTX model, forward pass and compression/decompression methods.
+Nardos Estifanos was responsible for evaluation of the results.
+All of us were partially responsible for the report for our corresponding parts.
 
@@
 
-\pagebreak
-
 \begin{figure*}[h!]
@@
 \vfill\pagebreak
+
+Like seen from the Fig.
+2 our models performance in quantitative evaluation is worse than any conventional method apart from the JPEG in very low bitrates.
+The BD-SNR is -8.32 dB and BD-rate 3.84\% compared to JPEG.
+In calculations we used only JPEG compressors those \textit{bdd} were in the same range than our models.
+The comparison is not fear because the \textit{bdd} ranges are slightly different as discussed above.
+From our models, we can compare the performance of two models that were trained on $\lambda$=8e-4 by using the KODAK dataset \cite{kodak}.
+Other models are not comparable in with BD-rates because the range of \textit{bbp} are different.
+In terms of BD-rate we got 3.987\% from the model that was trained only 250 epochs and N=M=192 (Model 1) compared to the one trained for 300 epochs with N=128 and M=192 (Model 3) with yields that less trained model with more parameters performs better.
+This is reasonable since bigger model should have better quality even with bit less training.
 

### 2023-03-20 paper/paper.tex
Add github repository link + fix formatting.

@@
 \usepackage{float}
+
+\usepackage{hyperref}
 
@@
 \end{abstract}
-%
-\begin{keywords}
-Learned image compression, ELIC Efficient Learned Image Compression
-\end{keywords}
-%
 
@@
 
-Juha Ylikoski was responsible for creating the cnns around the SCCTX-model and training the networks.
-Teemu Eerola was responsible for creating the SCCTX model, forward pass and compression/decompression methods.
+Juha Ylikoski was responsible for creating the cnns around the SCCTX and training the networks. Teemu Eerola was responsible for creating the SCCTX, forward pass and compression/decompression methods.
 Nardos Estifanos was responsible for evaluation of the results.
@@
 This work tries to replicate paper ELIC: Efficient learned image compression with unevenly grouped space-channel contextual adaptive coding \cite{ELIC} which has introduced an efficient learned image compression method (ELIC) which fairs well in the benchmarks and is also reasonable in terms of computational cost.
+We have created a github repository which is freely available \footnote{github repository \url{https://github.com/juha-ylikoski/learned_image_compression}}.
 The model is based on multiple convolutional neural networks which encode the image into latent space.
